Incentives are the biggest barriers to widespread data sharing

### 2021-08-13 new-paper.Rmd
small improvements to text of section 1 completed.

```{r incentable-table, fig.cap = "Table of the mechanism for behavior change, the incentive, and our assessment of where the elements of Data, Code, and Computational Environment rank in terms of completing these aspects. We note that Data is often required, but the preceding steps are not required. This is in contrast to Code having no policy."}
knitr::include_graphics("figures/incentives_pyramid.png")
```

Nosek [@nosek_2021] describes the strategies for culture change, especially when the status quo is too entrenched. Beyond motivation, skills, tools, and incentives are critical to enable such change. In a five step process, Nosek describes how achieving lasting culture change requires us to show proof of concept that an activity is possible, then improve the infrastructure to make it easy, followed by increased awareness to make it normative. Once there is a critical mass, find ways to reward those who put it into practice. After these steps, it would then be possible to require this of everyone (Figure \@ref(fig:incentable-table)). This framework provides a useful lens through which to examine adoption of code, computational environments, and data for reproducibility. With code, we have slowly progressed through these stages. While few journals require that authors share code, the venues to share, along with the necessary training have paved the way for incentives and credit. All of the necessary foundation to make code sharing required are in place. For computational environments, Docker makes it possible to share them in most common use cases [@Boettiger2015]. But running them still requires expertise and time on the part of reviewers. More recent open services like Binder [@Jupyter2018] make it possible to run notebooks atop Docker images, but the technology is still new and experiencing growing pains. There is still a long way to go before sharing computational environments in academia becomes mainstream. Finally, for data, we have implemented these steps out of order. The FAIR data principles were introduced in 2016 with good intentions [@Wilkinson2016] and describe the desired attributes for research data. That they be findable (good metadata), accessible (deposited somewhere persistent), interoperable (easily combined with other datasets), and be reusable. We as a community started out by making data required before demonstrating how this would directly benefit those who share data. Researchers are asked to deposit their data in one of many institutional or commercial repositories. Many journals even require it as a condition for publication via policy. However, the cost of preparing data for reuse remains high, and the career incentives to do so are almost non-existent. Until we make data reuse easier and more rewarding, no level of policy is going to improve the state of data sharing.

While there has ben a growing trend of sharing data alongside publications, the status quo of data sharing in science is not good. Rowhani-Farid et al. [@Rowhani-Farid2016] showed that 7 of 157 (4.5%) sampled papers in the British Medical Journal shared data, and only 14% of articles claiming open sharing of data actually shared their data. Similarly, Stodden et al. [@Stodden2018] showed only 36% of papers in the journal *Science* shared data, despite the Journal's data sharing policy. Similar findings have been found for PLOS and BMC journals [@Colavizza2020]. There are numerous examples of similar problems in other fields, such as psychology [@vanpaemel2015] and biology [@schofield2009]. The reality is that while data sharing is increasing in many fields such as ecology and genomics, the amount of data that can be reused is likely lower. To make shared data useful, we need to realign incentives for researchers.

Tools to prepare datasets and metadata (*the infrastructure*) and minimal set of practices (*guidelines*) for reuse are sparse [@Arslan2019; @Roche2015]. Unlike code, datasets are far more diverse in terms of heterogeneity, size, and formats. This makes them particularly challenging to standardize or easily "install" where the code is running. There are some efforts working towards improving this process, such as gogetdata , the 'dat' project, and a private company, Quilt.

Depositing data with rich metadata is laborious making it a pro forma activity outside of speciality repositories like Dryad that employ curators. Lastly, the infrastructure to track and get credit for data is non-existent. Datasets currently do not accrue citation credit since they are not tracked [@lowenberg_2019].

>>>>

```{r incentable-table, fig.cap = "Table of the mechanisms for behavior change, the incentives, and our assessment of where the elements of data, code, and computational environment rank in terms of completing these aspects. We note that data is often required, but the preceding steps are not, in contrast to code, which has no policy."}
knitr::include_graphics("figures/incentives_pyramid.png")
```

Nosek [@nosek_2021] describes the strategies for culture change, especially when the status quo is too entrenched. For change to occur, more than just motivation is needed: Skills, tools, and invenctives are critical. In a five-step process, Nosek describes how achieving lasting culture change requires us to show proof of concept that an activity is possible, improve the infrastructure to make it easy, and then increase awareness to make it normative. Once there is a critical mass, ways must be found to reward those who put it into practice. Once change has become normative and is widely practices, it is possible to require it of everyone (Figure \@ref(fig:incentable-table)). This framework provides a useful lens through which to examine adoption of code, computational environments, and data for reproducibility. With code, we have slowly progressed through these stages. While few journals require that authors share code, the provision of venues within which code can be shared, along with the necessary training, have paved the way for incentives and credit. All of the necessary foundations to make code sharing required are in place. For computational environments, Docker makes it possible to share them in most common use cases [@Boettiger2015]. But running them still requires expertise and time on the part of reviewers. More recent open services, like Binder [@Jupyter2018], make it possible to run notebooks atop Docker images, but the technology is still new and experiencing growing pains. There is still a long way to go before sharing computational environments in academia becomes mainstream.

Finally, for data, we have implemented these steps out of order. The FAIR data principles, introduced in 2016 with good intentions, [@Wilkinson2016] describe the desired attributes for research data: that they be findable (good metadata), accessible (deposited somewhere persistent), interoperable (easily combined with other datasets), and reusable. We, as a community, started out by making data required before demonstrating how this would directly benefit those who share data. Researchers are asked to deposit their data in one of many institutional or commercial repositories. Many journals even require it as a condition for publication via policy. However, the cost of preparing data for reuse remains high, and career incentives to do so are almost nonexistent. Until we make data reuse easier and more rewarding, no level of policy is going to improve the state of data sharing.

While there has been a growing trend of sharing data alongside publications, the status quo of data sharing in science is not good. Rowhani-Farid et al. [@Rowhani-Farid2016] showed that 7 of 157 (4.5%) sampled papers in the *British Medical Journal* shared data, and only 14% of articles claiming open sharing of data actually shared their data. Similarly, Stodden et al. [@Stodden2018] showed that only 36% of papers published in the journal *Science* shared data, despite the journal's data sharing policy. Similar findings have been made for PLOS and BMC journals [@Colavizza2020]. There are numerous examples of similar problems in other fields, such as psychology [@vanpaemel2015] and biology [@schofield2009]. The reality is that while data sharing is increasing in many fields, such as ecology and genomics, the amount of data that can be reused is likely lower. To make shared data useful, we need to realign incentives for researchers.

Tools that can be used to prepare datasets and metadata (*the infrastructure*) and minimal sets of practices (*guidelines*) for reuse are sparse [@Arslan2019; @Roche2015]. Unlike code, datasets are far more diverse in terms of heterogeneity, size, and format. This makes them particularly challenging to standardize or easily "install" where the code is running. There have been some efforts to improve this process, such as gogetdata, the Dat Project, and the formation of a private company, Quilt.

Depositing data with rich metadata is laborious, making it a pro forma activity, outside of speciality repositories like Dryad that employ curators. Finally, the infrastructure to track and get credit for data is does not yet exist. Datasets currently do not accrue citation credit since they are not tracked [@lowenberg_2019].
